Activity data analysis
========================================================
* Loading and preprocessing the data
read the data set.
```{r, echo = TRUE}
activity <- read.csv(unz("RepData_PeerAssessment1//activity.zip", "activity.csv"), header = T)
#dim(activity)
#head(activity)
```
* what is mean total number of steps taken per day?

draw the histogram of total steps taken of each day
```{r echo = TRUE}
stepsOfEachDay <- as.data.frame(sapply(split(activity$steps,activity$date), sum))
names(stepsOfEachDay) <-"steps"
hist(stepsOfEachDay$steps, xlab = "total steps of day")
```

the mean of total number of steps taken per day
```{r echo = TRUE}
mean(stepsOfEachDay$steps, na.rm = TRUE)
```

the median of total number of steps taken per day
```{r echo = TRUE}
median(stepsOfEachDay$steps, na.rm = TRUE)
```

* What is the average daily activity partten?
summerize the activity according to the time interval
```{r echo = TRUE}
stepOfEachInterval <- as.data.frame(sapply(split(activity$steps,activity$interval), mean, na.rm = TRUE))
names(stepOfEachInterval) <- "steps"
head(stepOfEachInterval)
plot(stepOfEachInterval$steps, type = "l", ylab = "average steps", xlab = "intervals", main = "average steps at each interval")
```
find out which interval contains the maximum number of steps
```{r, echo = TRUE}
stepOfEachInterval$interval <- as.numeric(as.character(rownames(stepOfEachInterval)))
thisInterval <-  which.max(stepOfEachInterval$steps)
### the maximum number of steps
stepOfEachInterval$steps[thisInterval]
### this interval has the maximum number of steps
stepOfEachInterval$interval[thisInterval]
```

* imputing missing values
calculate the number of NA
```{r, echo = TRUE}
numberOfNA <- is.na(activity$steps) 
table(numberOfNA)
```

filling in the missing value with the average steps at this interval
```{r, echo = TRUE}
### make copy of the activity
activityModified <- activity
### merge the two data sets according to interval column 
activityM <- merge(activityModified, stepOfEachInterval, by = "interval", suffixes = c(".activityModified", ".stepOfEachInterval"))

### substitute the NA in the activityModified data set according to activityM
activityModified$steps = ifelse(is.na(activityM$steps.activityModified), activityM$steps.stepOfEachInterval, activityM$steps.activityModified)

head(activityModified)
```
draw the histogram of total steps taken of each day
```{r echo = TRUE}
stepsOfEachDayModified <- as.data.frame(sapply(split(activityModified$steps,activityModified$date), sum))
names(stepsOfEachDayModified) <-"steps"
hist(stepsOfEachDayModified$steps)
```


the mean of total number of steps taken per day for modified activity
```{r echo = TRUE}
mean(activityModified$steps)
```

the median of total number of steps taken per day for modified activity
```{r echo = TRUE}
median(activityModified$steps)
```
average number of steps is much smaller than before after filling in the missing value with the average, therefore there is no effect.

* Are there difference in activity patterns between weekday and weekends?
```{r, echo = TRUE}
activityModified$date <- as.Date(activityModified$date)
activityModified$day <- weekdays(activityModified$date)
head(activityModified)
activityModified$type <- ifelse((activityModified$day =="Sunday") |(activityModified$day =="Saturday"), "weekend", "weekday") ### convert day to weekend and weekday 
```

```{r, echo = TRUE}
library(ggplot2)
### split data to weekday and weekend
activityweekday <- subset(activityModified, type == "weekday")
activityweekend <- subset(activityModified, type == "weekend")
###calculate the weekday average steps cross all weekdays for each interval
weekdayMeanStep <- as.data.frame(sapply(split(activityweekday$steps, activityweekday$interval), mean))
colnames(weekdayMeanStep) <- "steps"
nrow(weekdayMeanStep)
weekdayMeanStep$type <- rep("weekday", 288)
weekdayMeanStep$interval <- as.numeric(as.character(rownames(weekdayMeanStep)))

### calculate the weekend average steps cross weekends for each interval
weekendMeanStep <- as.data.frame(sapply(split(activityweekend$steps, activityweekend$interval), mean))
colnames(weekendMeanStep) <- "steps"
nrow(weekendMeanStep)
weekendMeanStep$type <- rep("weekend", 288)
weekendMeanStep$interval <- as.numeric(as.character(rownames(weekendMeanStep)))

### rbind weekday and weekend data
meanStepCombined <- rbind(weekdayMeanStep, weekendMeanStep)
#dim(meanStepCombined)
#class(meanStepCombined)
#head(meanStepCombined)
#tail(meanStepCombined)

###graph data
p <- ggplot(meanStepCombined, aes(x = interval, y = steps, group = type))
p + geom_line(aes(group = type, color = type)) + ylab("number of steps")
```